---
title: "Math 408"
subtitle: "Machine Learning"
format:
  revealjs:
    width: 1200
    scrollable: false
    footer: m408.inqs.info/lectures/1a
    theme: [default, styles.scss]
    preview-links: true
    navigation-mode: vertical
    controls-layout: edges
    controls-tutorial: true
    slide-number: true
    pointer:
      pointerSize: 32
    incremental: false 
    chalkboard:
      theme: whiteboard
      chalk-width: 4
knitr:
  opts_chunk: 
    echo: true
    eval: true
    message: false
    code-fold: true
    warnings: false
    comment: "#>" 
    
revealjs-plugins:
  - pointer
  - verticator
  
filters: 
  - code-fullscreen
  - reveal-auto-agenda

editor: source
---

# Installation

## R Programming

R is a statistical programming package that allows you to conduct different types of analysis.

[R](https://www.r-project.org/)

::: callout-note
Windows Users: Install [RTools](https://cran.r-project.org/bin/windows/Rtools/)
:::

## Positron

A piece of software that organizes how you conduct statistical analysis in R.

[Positron](http://positron.posit.co/download.html)

::: callout-note
Windows Users: Use system-level package
If you cannot work with R, install [Visual C++ Redistributable](https://learn.microsoft.com/en-us/cpp/windows/latest-supported-vc-redist?view=msvc-170#latest-microsoft-visual-c-redistributable-version)
:::

## RStudio

If Positron does not work, install [RStudio](https://posit.co/downloads/) instead.

## Installing Torch

In R, install torch with:

```r
install.packages("torch")
```

::: callout-note

Nvidia-Users

```r
options(timeout = 600) # increasing timeout is recommended since we will be downloading a 2GB file.
# For Windows and Linux: "cpu", "cu128" are the only currently supported
# For MacOS the supported are: "cpu-intel" or "cpu-m1"
kind <- "cpu"
version <- available.packages()["torch","Version"]
options(repos = c(
  torch = sprintf("https://torch-cdn.mlverse.org/packages/%s/%s/", kind, version),
  CRAN = "https://cloud.r-project.org" # or any other from which you want to install the other R dependencies.
))
install.packages("torch")
```
:::

## Installing Stan

```r
install.packages("cmdstanr", 
    repos = c('https://stan-dev.r-universe.dev', getOption("repos")))
cmdstanr::check_cmdstan_toolchain()
cmdstanr::install_cmdstan(cores = 2)
```

## Other R Packages


:::: panel-tabset

### Core Packages

-   Tidyverse


### Course Packages


:::: columns

::: column

- ISLR2

- leaps

- gam

- tree


:::

::: column

- gbm

- e1071

- glmnet

- torch

:::

::::

### Misc Packages

- RMarkdown

- Quarto

- extraInserts (GitHub Package)

::::

## Load R Package

```r
library(tidyverse)
library(torch)
```


- You must load packages every new R Session

# Workspaces

# R Basics

## R as a calculator

R can evaluate different expressions in the console tab.

Try the following:

1.  $4(4+2)/34$
2.  $6^3$
3.  $3-1$
4.  $4+4/3+45(32*34-54)$


## R Functions

R functions performs tasks to specific data values.

Evaluate the following values in R:

1.  $\sqrt{3}$
2.  $e^3$
3.  $\ln(53)$
4.  $\log(324)$
5.  $\sin(3)$
6.  $\sin(3\pi)$

## Types of Data

-   Numeric

-   Character

-   Logical

-   Missing

Evaluate the following code:

```r
is.numeric(1)
is.numeric("1")
is.numeric(T)
is.numeric(NA)

```

## Types of Objects

In R, an object contains a set of data. The most common types are `vectors` and `matrix`.

Run this code and print out the objects in the console:

```r
x <- 3:34
y <- matrix(1:20, nrow = 4)
```


## Data Frames

Data frames can be thought of as R's version of a data set.

Play around with mtcars:

```{r}
#| eval: true
mtcars 
```

## Lists

List can be thought as an extended vector, but each element is a different R object.

Try playing with this R object:

```{r}
list_one <- list(mtcars, rep(0, 4),
                 diag(rep(1, 3)))
list_one
```

