---
title: "R Basics"
subtitle: "Functions"
format:
  revealjs:
    width: 1200
    scrollable: false
    include-in-header: "math.html"
    footer: m408.inqs.info/lectures/1a
    theme: [default, styles.scss]
    preview-links: true
    navigation-mode: vertical
    controls-layout: edges
    controls-tutorial: true
    slide-number: true
    pointer:
      pointerSize: 48
    incremental: false 
    chalkboard:
      theme: whiteboard
      chalk-width: 4
knitr:
  opts_chunk: 
    echo: true
    eval: true
    message: false
    warnings: false
    comment: "#>" 
    
revealjs-plugins:
  - pointer
  - verticator
  
filters: 
  - code-fullscreen
  - reveal-auto-agenda

editor: source
---

## R Packages

```{r}
library(tidyverse)
```

# Indexing

## Indexing

Within an R object, you can access an element by indexing it.

Indexing tells R which values to output.

## Vectors

A vector can be indexed by adding `[]` after the object's name and specifying the number of each element.

```{webr-r}
letters
letters[13]
```

## Matrices

A matrix can be indexed by adding `[]` after the object's name and specifying the number of each element. Separate the values by commas for specific indexes.

```{webr-r}
x <- matrix(1:40, nrow = 4)
```

## Data Frames

Data frames can be indexed using the `$` operator and `[]`.

```{webr-r}
mtcars[,"mpg"]
```

## Lists

Lists can be indexed using the `[[]]` for a specific element of a list.

```{webr-r}
toy_list <- list(x = letters,
                 y = mtcars,
                 z = list(x = diag(rep(1, 5),
                          y = matrix(1:40, nrow = 5),
                          z = band_members)))
```

# `if`/`else` Statements

## if/else Statements

if/else statements are used to conduct specific tasks depending on the conditions

## `if` Statement

An `if` statement is used to if you want R to perform a specific function if a certain condition is met. An `if` statement will only run a task if a logical is returned. You will need type `if`, followed by the condition (as a logical) in parentheses, then the task.

## Example

```{webr-r}
x <- sample(-10:10,1)
if (x > 0){
  print("Positive")
}
print(x)

```

## else statement

An `else` statement will conduct a different task if the `if` statement does not conduct the tasks.

## Example

```{webr-r}
x <- sample(-10:10,1)
if (x > 0 ){
  print("Positive")
} else {
  print("Non-positive")
} 
print(x)

```

## Chain if/else statement

If you have more than two options, you can chain `if/else` statements by adding an `if` statement immediately after the word `else`.

## Example

```{webr-r}
x <- sample(-10:10,1)
if (x > 0 ){
  print("Positive")
} else if (x == 0) {
  print("Zero")
} else {
  print("Negative")
}
print(x)

```

# Built-in Functions

## Built-in Functions

There are several available functions in R to conduct specific statistical methods or tasks

## Help Documentation

| Section     | Description                                                 |
|:-------------------|:---------------------------------------------------|
| Description | Provides a brief introduction of the function               |
| Usage       | Provides potential usage of the function                    |
| Arguments   | Arguments that the function can take                        |
| Details     | An in depth description of the function                     |
| Value       | Provides information of the output produced by the function |
| Notes       | Any need to know information about the function             |
| Authors     | Developers of the function                                  |
| References  | References to the model and function                        |
| See Also    | Provide information of supporting functions                 |
| Examples    | Examples of the function                                    |

## Generic Functions

Several R objects have a known class attached to it. A specialized object designed to be read by generic functions, such as `summary()` and `plot()`.

For example, the `summary()` is a generic for several types of functions: `summary.aov()`, `summary.lm()`, `summary.glm()`, and many more.

## Commonly-used Function

| Functions  | Description                 |
|:-----------|:----------------------------|
| `aov()`    | Fits an ANOVA Model         |
| `lm()`     | Fits a linear model         |
| `glm()`    | Fits a general linear model |
| `t.test()` | Conducts a t-test           |

# User-built functions

## User-built functions

-   Functions created by the user for analysis

-   Needs to be ran once to the R environment

-   Will be lost when R session is closed

## Anatomy

```{r}
#| eval: false
#| code-line-numbers: "|1-3|4-5|6|7"

name_of_function <- function(data_1, data_2 = NULL, 
                             argument_1, argument_2 = TRUE, argument_3 = NULL,
                             ...){
  # Conduct Task
  # Conduct Task
  output_object <- Tasks
  return(output_object)
}

```

::: fragment
-   `function`: used to construct the function

-   `data1`: first data argument that needs to supplied

-   `data2`: second data argument that does not need to be supplied

-   `argument1`: first argument must be supplied to alter function

-   `argument2`: second argument to alter function, set to `TRUE`

-   `argument3`: third argument that does not need to be supplied

-   `…`: additional arguments supplied to other functions
:::

## Example

Create a function for

$$
y = \ln(x^2)
$$

## Example

Create a function for

$$
f(x) = \left\{\begin{array}{cc}
x^3 & x<0\\
x^2 + 5 & \mathrm{otherwise}
\end{array} \right.
$$

## Example

Create a function for

$$
f(x,y) = \left\{\begin{array}{cc}
x^3 e^y &  x<0\ \\
x^2 + 5 + \ln(y) & \mathrm{otherwise}
\end{array} \right.
$$

## Example

Create the function that allows your to compute the z-score of a specific value `x` using the sampling distribution from a set of data (`y` vector):

$$
z =  \frac{x-\bar y}{\sqrt{s^2_{y}/n_y}}
$$
