---
title: "R Basics"
subtitle: "Functions"
format:
  revealjs:
    width: 1200
    scrollable: false
    include-in-header: "math.html"
    footer: m408.inqs.info/lectures/1a
    theme: [default, styles.scss]
    preview-links: true
    navigation-mode: vertical
    controls-layout: edges
    controls-tutorial: true
    slide-number: true
    pointer:
      pointerSize: 48
    incremental: false 
    chalkboard:
      theme: whiteboard
      chalk-width: 4
knitr:
  opts_chunk: 
    echo: true
    eval: true
    message: false
    warnings: false
    comment: "#>" 
    
revealjs-plugins:
  - pointer
  - verticator
  
filters: 
  - code-fullscreen
  - reveal-auto-agenda

editor: source
---

## R Packages

```{r}
library(tidyverse)
```

# Class Setup

## Class Setup Modifications 

::: {.incremental}

- Tuesday
  - Lecture on Neural Networks
- Thursday
  - Work on Project
  - Work on Goals that leads to the completion of your project.
:::

::: fragment
This will begin next week.
:::

## Why the change

This will help ensure projects get done, and students to learn at their own pace. 

## Goals

Every week, your homework is to create one SMART Goal and provide evidence that you reached that goal.

::: fragment

If you were not able to reach your goal, provide clear steps on how to achieve it the following week.

:::

## SMART Goals

SMART goals are a structured way to set clear, achievable objectives.

They help turn vague intentions into concrete plans.


## What Does SMART Stand For?

- **S**pecific: Be clear and focused.
- **M**easurable: A criteria to track progress.
- **A**chievable: Be realistic and attainable.
- **R**elevant: Make it align with your priorities.
- **T**ime-bound: Have a clear deadline.



## Example (Not SMART)

> I want to get better at R programming.


## Example (SMART)

> I will learn how to use `for` loops in R by writing at least one loop-based script in a week that performs tasks such as iterating over vectors and data frames.

## Why Use SMART Goals?

* Improves focus and motivation
* Makes progress measurable
* Increases likelihood of success
* Reduces ambiguity

## Homework 0

Provide some background information about your experience in computing. Have you coded any algorithms? Have you've done any statistical analysis? What programming languages have used before? 

Write a paragraph of what you want to achieve in this machine learning course, and create one smart goal to achieve by next Thursday.

Provide a description of what you want to do for your project.



# Workspaces

## Workspaces

Workspaces allows you to create locations on your computer and set up a computing environment in one location.

# R Basics

## R as a calculator

R can evaluate different expressions in the console tab.

Try the following:

1.  $4(4+2)/34$
2.  $6^3$
3.  $3-1$
4.  $4+4/3+45(32*34-54)$


## R Functions

R functions performs tasks to specific data values.

Evaluate the following values in R:

1.  $\sqrt{3}$
2.  $e^3$
3.  $\ln(53)$
4.  $\log(324)$
5.  $\sin(3)$
6.  $\sin(3\pi)$

## Types of Data

-   Numeric

-   Character

-   Logical

-   Missing

Evaluate the following code:

```r
is.numeric(1)
is.numeric("1")
is.numeric(T)
is.numeric(NA)

```

## Types of Objects

In R, an object contains a set of data. The most common types are `vectors` and `matrix`.

Run this code and print out the objects in the console:

```r
x <- 3:34
y <- matrix(1:20, nrow = 4)
```


## Data Frames

Data frames can be thought of as R's version of a data set.

Play around with mtcars:

```{r}
#| eval: true
mtcars 
```

## Lists

List can be thought as an extended vector, but each element is a different R object.

Try playing with this R object:

```{r}
list_one <- list(mtcars, rep(0, 4),
                 diag(rep(1, 3)))
list_one
```

# Comparing Numbers

## Comparing Numbers

You can compare two numbers, or objects, that will result in a logical output.

## Comparing Numbers Operators

| Operator | Description           |
|----------|-----------------------|
| `>`      | Greater Than          |
| `<`      | Less Than             |
| `>=`     | Greater than or equal |
| `<=`     | Less than or equal    |
| `==`     | Equals                |
| `!=`     | Not Equals            |

## Labeled Lists (Key-Value)

```{r}
list_two <- list(data = mtcars, 
                 vector  = rep(0, 4),
                 matricx = diag(rep(1, 3)))
list_two

list_two$vector
```

## Resources in Programming

[R Basics Material](https://www.inqs.info/stat_comp/r.html)

# Indexing

## Indexing

Within an R object, you can access an element by indexing it.

Indexing tells R which values to output.

## Vectors

A vector can be indexed by adding `[]` after the object's name and specifying the number of each element.

```r
letters
letters[13]
```

## Matrices

A matrix can be indexed by adding `[]` after the object's name and specifying the number of each element. Separate the values by commas for specific indexes.

```r
x <- matrix(1:40, nrow = 4)
```

## Data Frames

Data frames can be indexed using the `$` operator and `[]`.

```r
mtcars[,"mpg"]
```

## Lists

Lists can be indexed using the `[[]]` for a specific element of a list.

```r
toy_list <- list(x = letters,
                 y = mtcars,
                 z = list(x = diag(rep(1, 5),
                          y = matrix(1:40, nrow = 5),
                          z = band_members)))
```

# `if`/`else` Statements

## if/else Statements

if/else statements are used to conduct specific tasks depending on the conditions

## `if` Statement

An `if` statement is used to if you want R to perform a specific function if a certain condition is met. An `if` statement will only run a task if a logical is returned. You will need type `if`, followed by the condition (as a logical) in parentheses, then the task.

## Example

```r
x <- sample(-10:10,1)
if (x > 0){
  print("Positive")
}
print(x)

```

## else statement

An `else` statement will conduct a different task if the `if` statement does not conduct the tasks.

## Example

```r
x <- sample(-10:10,1)
if (x > 0 ){
  print("Positive")
} else {
  print("Non-positive")
} 
print(x)

```

## Chain if/else statement

If you have more than two options, you can chain `if/else` statements by adding an `if` statement immediately after the word `else`.

## Example

```r
x <- sample(-10:10,1)
if (x > 0 ){
  print("Positive")
} else if (x == 0) {
  print("Zero")
} else {
  print("Negative")
}
print(x)

```

# Built-in Functions

## Built-in Functions

There are several available functions in R to conduct specific statistical methods or tasks

## Help Documentation

| Section     | Description                                                 |
|:-------------------|:---------------------------------------------------|
| Description | Provides a brief introduction of the function               |
| Usage       | Provides potential usage of the function                    |
| Arguments   | Arguments that the function can take                        |
| Details     | An in depth description of the function                     |
| Value       | Provides information of the output produced by the function |
| Notes       | Any need to know information about the function             |
| Authors     | Developers of the function                                  |
| References  | References to the model and function                        |
| See Also    | Provide information of supporting functions                 |
| Examples    | Examples of the function                                    |

## Generic Functions

Several R objects have a known class attached to it. A specialized object designed to be read by generic functions, such as `summary()` and `plot()`.

For example, the `summary()` is a generic for several types of functions: `summary.aov()`, `summary.lm()`, `summary.glm()`, and many more.

## Commonly-used Function

| Functions  | Description                 |
|:-----------|:----------------------------|
| `aov()`    | Fits an ANOVA Model         |
| `lm()`     | Fits a linear model         |
| `glm()`    | Fits a general linear model |
| `t.test()` | Conducts a t-test           |

# User-built functions

## User-built functions

-   Functions created by the user for analysis

-   Needs to be ran once to the R environment

-   Will be lost when R session is closed

## Anatomy

```{r}
#| eval: false
#| code-line-numbers: "|1-3|4-5|6|7"

name_of_function <- function(data_1, data_2 = NULL, 
                             argument_1, argument_2 = TRUE, argument_3 = NULL,
                             ...){
  # Conduct Task
  # Conduct Task
  output_object <- Tasks
  return(output_object)
}

```

::: fragment
-   `function`: used to construct the function

-   `data1`: first data argument that needs to supplied

-   `data2`: second data argument that does not need to be supplied

-   `argument1`: first argument must be supplied to alter function

-   `argument2`: second argument to alter function, set to `TRUE`

-   `argument3`: third argument that does not need to be supplied

-   `…`: additional arguments supplied to other functions
:::

## Example

Create a function for

$$
y = \ln(x^2)
$$

## Example

Create a function for

$$
f(x) = \left\{\begin{array}{cc}
x^3 & x<0\\
x^2 + 5 & \mathrm{otherwise}
\end{array} \right.
$$

## Example

Create a function for

$$
f(x,y) = \left\{\begin{array}{cc}
x^3 e^y &  x<0\ \\
x^2 + 5 + \ln(y) & \mathrm{otherwise}
\end{array} \right.
$$

## Example

Create the function that allows your to compute the z-score of a specific value `x` using the sampling distribution from a set of data (`y` vector):

$$
z =  \frac{x-\bar y}{\sqrt{s^2_{y}/n_y}}
$$
